---
title: "Data Wrangling and Analysis of Experiment Results"
author: "Masoud Jasbi"
date: "7/8/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

```{r libraries, include=FALSE}
library(tidyverse)
library(ggthemes)
library(VIM)
```

## Data Wrangling

```{r import_data}
qualtrics_data <- read.csv("data/raw/Pilot-Quantifiers-data-trimmed.csv", na.strings = "")

N_dropout <-
  qualtrics_data %>%
  filter(Finished=="FALSE") %>%
  nrow()

N_participants <- 
    qualtrics_data %>%
  filter(Finished=="TRUE") %>%
  nrow()
```

Total of `r N_dropout + N_participants` took this survey. `r N_dropout` did not finish the survey and dropped out. The results for `r N_participants` are reported.

```{r}
tidy_data <-
  qualtrics_data %>%
  filter(Finished=="TRUE") %>%
  mutate(sid=c(1:56)) %>%
  select(-Finished) %>%
  gather(quantifier, response, Every_day:Never) %>%
  drop_na(response) %>%
  mutate(condition = c(rep("nominal", 255), rep("temporal", 297))) %>%
  mutate(response = as.numeric(response))

write_csv(tidy_data, "data/processed/Pilot-Quantifiers-data-processed.csv")
```

## Plots

```{r QuantifierDistribution}
tidy_data %>%
  ggplot(aes(response)) +
  geom_histogram() + 
  facet_wrap(.~quantifier) +
  theme_few()
```

```{r conditionDistribution}
tidy_data %>%
  ggplot(aes(response)) +
  geom_histogram() + 
  facet_wrap(.~condition) +
  theme_few()
```


```{r boxplot}
tidy_data$quantifier <- reorder(tidy_data$quantifier, tidy_data$response, mean)

tidy_data %>%
  ggplot(aes(quantifier, response, color=condition), fill=condition) +
  geom_boxplot() +
  geom_jitter(size=1) +
  theme_few() + 
  theme(axis.text.x = element_text(angle=45, hjust = 1, vjust = 1))
```

## Analysis

K-means Clustering

```{r kmeansClustering}
kmeans_data1 <-
  tidy_data %>%
  filter(condition=="nominal") %>%
  spread(quantifier, response) %>%
  select(-sid, -condition)

kmeans_data2 <-
  tidy_data %>%
  filter(condition=="temporal") %>%
  spread(quantifier, response) %>%
  select(-sid, -condition)

kmeans_data <- 
  bind_cols(kmeans_data1, kmeans_data2) %>%
  select(-no_days, Never) %>%
  drop_na()

df2 <- data.frame(t(kmeans_data[-1]))
colnames(df2) <- kmeans_data[, 1]
  
quantifier_clusters <- kmeans(df2, centers=4, nstart = 25)
quantifier_clusters$cluster
```

4 clusters:

* each-day, every-day, all-days, always
* often, usually, frequently, mostly, many-days, most-days
* infrequently, sometimes, some-days,
* few-days, several-days, seldom, rarely, occasionally, (never)

5 clusters:

* every-day, each-day, always
* often, usually
* most-days, many-days, frequently, mostly
* few-days, several-days, rarely, occasionally, never
* some-days, some-times
