---
title: "Data Wrangling and Analysis of Experiment Results"
author: "Masoud Jasbi"
date: "7/8/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

```{r libraries, include=FALSE}
library(tidyverse)
library(ggthemes)
library(VIM)
library(cluster)    # clustering algorithms
library(factoextra) # clustering visualization
library(dendextend)
```

## Data Wrangling

```{r import_data}
qualtrics_data <- read.csv("data/raw/Pilot-Quantifiers-data-trimmed.csv", na.strings = "")

N_dropout <-
  qualtrics_data %>%
  filter(Finished=="FALSE") %>%
  nrow()

N_participants <- 
    qualtrics_data %>%
  filter(Finished=="TRUE") %>%
  nrow()
```

Total of `r N_dropout + N_participants` took this survey. `r N_dropout` did not finish the survey and dropped out. The results for `r N_participants` are reported.

```{r}
tidy_data <-
  qualtrics_data %>%
  filter(Finished=="TRUE") %>%
  mutate(sid=c(1:56)) %>%
  select(-Finished) %>%
  gather(quantifier, response, Every_day:Never) %>%
  drop_na(response) %>%
  mutate(condition = c(rep("nominal", 255), rep("temporal", 297))) %>%
  mutate(response = as.numeric(response))

write_csv(tidy_data, "data/processed/Pilot-Quantifiers-data-processed.csv")
```

## Plots

```{r QuantifierDistribution}
tidy_data %>%
  ggplot(aes(response)) +
  geom_histogram() + 
  facet_wrap(.~quantifier) +
  theme_few()
```

```{r conditionDistribution}
tidy_data %>%
  ggplot(aes(response)) +
  geom_histogram() + 
  facet_wrap(.~condition) +
  theme_few()
```


```{r boxplot}
tidy_data$quantifier <- reorder(tidy_data$quantifier, tidy_data$response, mean)

tidy_data %>%
  ggplot(aes(quantifier, response, color=condition), fill=condition) +
  geom_boxplot() +
  geom_jitter(size=1) +
  theme_few() + 
  theme(axis.text.x = element_text(angle=45, hjust = 1, vjust = 1))
```


## Data Wrangling

```{r}
# let's retrieve the question quantifier and context from the handcoded dataframe
handTrimmed <- read.csv("data/handtrimmed-final-quantifiers.csv")

columnVariables <- 
  handTrimmed %>%
  select(-mTurkCode, -Age, -Language, -UserLanguage) %>%
  mutate(sid=c(1:nrow(handTrimmed))) %>%
  colnames()
```

```{r import_data}
qualtrics_data <- read.csv("data/quantifiers_data_raw.csv", na.strings = "")

#let's remove the previews and the ones we ran for testing ourselves
qualtrics_data <- qualtrics_data[-1:-58,]
```

```{r exclusions}
# How many people dropped out of the experiment?
N_dropout <-
  qualtrics_data %>%
  filter(Finished=="False") %>%
  nrow()

# How many people did not speak English?
N_nonEnglish <-
  qualtrics_data %>%
  filter(Q239=="46" | Q239=="Italian" | Q239=="Spanish")

# How man were under 18? (no one)
N_under18 <-
  qualtrics_data %>%
  filter(Q238=="Under 18")

# How many failed the attention checks?
N_inattentive <-
  qualtrics_data %>%
  filter(Q240_1!=70 | Q246_1!=30)

# How many people are included?
included_data <- 
  qualtrics_data %>%
  filter(Finished=="True") %>%
  filter(Q239!="46", Q239!="Italian", Q239!="Spanish") %>%
  filter(Q238!="Under 18") %>%
  filter(Q240_1==70 | Q246_1==30)

# number of included participants
N_participants <-
  included_data %>%
  nrow()

# check to see if number of IP addresses match number of participants
N_IPAddress <- included_data$IPAddress %>% unique() %>% length()
```

Total of `r nrow(qualtrics_data)` took this survey. `r N_dropout` did not finish the survey and dropped out. `r N_nonEnglish` reported their language as other than English. `r N_under18` participants were under 18. And `r N_inattentive` participants failed at least one of the attention checks. The results for `r N_participants` are reported.

```{r}
tidy_data <-
  included_data %>%
  mutate(sid=c(1:nrow(included_data))) %>% # add subject ID column
  select(-StartDate, -EndDate, -Status, -IPAddress, -Progress, -Duration..in.seconds.,-Finished, -RecordedDate, -ResponseId, -RecipientFirstName, -RecipientLastName, -RecipientEmail, -ExternalReference, -LocationLatitude, -LocationLongitude, -DistributionChannel, -UserLanguage, -mTurkCode, -Q238, -Q239) # remove useless columns column

det_every <- c("Every_GroceryStore", "Every_Early", "Every_Teeth", "Every_Jog
", "Every_Eggs", "Every_Drove", "Every_Sandwich", "Every_WorkLate", "Every_BedLate", "Every_Pizza", "Every_Biked")

tidy_data_ExtraExclusion <- 
  tidy_data %>%
  sapply(as.numeric) %>%
  as.data.frame() %>%
  filter_at(vars(contains("Every")), any_vars(.>5))

names(tidy_data) <- columnVariables

long_data <-
  tidy_data %>%
  gather(quantifier_context, response, Every_GroceryStore:Never_Biked) %>%
  drop_na(response) %>%
  separate(quantifier_context, c("quantifier", "context"), sep = "_")

# length(unique(long_data$sid)) 6 participants excluded because of sliders
# 

write.csv(long_data, "data/Quantifiers_processed_data.csv", row.names=FALSE)
```

## Plots

```{r response_plot}
long_data$response <- as.numeric(long_data$response)
long_data$quantifier <- reorder(long_data$quantifier, long_data$response, median)

long_data %>%
  filter(quantifier!="Slider") %>%
  ggplot(aes(quantifier, response)) +
  geom_boxplot(outlier.shape=NA) +
  geom_point(size=0.1) +
  geom_jitter(size=0.1) +
  stat_summary(fun.y=median, geom="point", shape=20, size=4, color="red", fill="red") +
  theme_classic() + 
  theme(axis.text.x = element_text(size= 16, angle=45, hjust = 1, vjust = 1)) + xlab("") + ylab("")
```

```{r}
long_data %>%
  filter(quantifier!="Slider") %>%
  ggplot(aes(quantifier, response)) +
  geom_jitter(size=0.5) +
  theme(axis.text.x = element_text(angle=45, hjust = 1, vjust = 1)) +
  theme_classic()

long_data %>%
  group_by(quantifier) %>%
  summarize(count=n())
```

# Hierarchical Cluster Analysis

```{r cluster_analysis}
cluster_data <-
  long_data %>%
  filter(quantifier!="Slider", quantifier!="Mostly", quantifier!="Often") %>%
  spread(quantifier, response) %>%
  select(-sid, -context)

hclust_data <- as.data.frame(lapply(cluster_data, na.omit))
length(new_plot$Always)
new_plot <-
  hclust_data %>%
  filter(No < 50, Every > 50, Never < 50, All > 50, Always >50, Each > 50) %>%
  nrow()
```


```{r}
cluster_data <-
  long_data %>%
  filter(quantifier!="Slider", quantifier!="Mostly", quantifier!="Often") %>%
  spread(quantifier, response) %>%
  select(-sid, -context)

hclust_data <- as.data.frame(lapply(cluster_data, na.omit))

cluster_data1 <-
  long_data %>%
  filter(quantifier=="Mostly") %>%
  spread(quantifier, response) %>%
  select(-sid, -context) %>%
  filter(Mostly!=0) %>%
  sample_n(97)

cluster_data2 <-
  long_data %>%
  filter(quantifier=="Often") %>%
  spread(quantifier, response) %>%
  select(-sid, -context) %>%
  sapply(as.numeric) %>%
  as.data.frame()

Often <- rep(median(cluster_data2$Often), 7)

cluster_data3 <- data.frame(Often)

addcolumns <- bind_rows(cluster_data2, cluster_data3)

hclust_data_final <- 
  bind_cols(hclust_data, cluster_data1, addcolumns) %>%
  rename(`A Few` = "AFew", Infrequently = "Infreq", Occasionally = "Occ", Frequently = "Freq")

write.csv(hclust_data_final, "data/hclustdata.csv")

hculst_data_final <- read.csv("data/hclustdata.csv", row.names = NULL)

finalplot <-
hculst_data_final %>%
  filter(No < 50, Every > 50, Never < 50, All > 50, Always >50, Each > 50) %>%
  select(-X) %>%
  rename(`A Few` = "A.Few")

newnew_plot <-
  finalplot %>%
  gather(quantifier, response)

#newnew_plot$response <- as.numeric(newnew_plot$response)
newnew_plot$quantifier <- reorder(newnew_plot$quantifier, newnew_plot$response, mean)

newnew_plot %>%
  spread (quantifier, response)

newnew_plot %>%
  ggplot(aes(quantifier, response)) +
  geom_boxplot(outlier.shape=NA) +
  geom_point(size=0.1) +
  geom_jitter(size=0.1) +
  stat_summary(fun.y=median, geom="point", shape=20, size=4, color="red", fill="red") +
  theme_classic() + 
  theme(axis.text.x = element_text(size= 16, angle=45, hjust = 1, vjust = 1)) + xlab("") + ylab("")

hclust_data_final_t <- data.frame(t(hclust_data_final))
colnames(hclust_data_final_t) <- hclust_data_final_t[, 1]

# Hierarchical clustering using Complete Linkage
hc1 <- hclust(
  dist(hclust_data_final_t, method = "manhattan"), 
  method = "complete")

# Plot the obtained dendrogram
plot(hc1, cex = 0.6, hang = 0.2)
```


