---
title: "Data Wrangling and Analysis of Experiment 2 Results"
author: "Masoud Jasbi"
date: "7/8/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

```{r libraries, include=FALSE}
library(tidyverse)
library(ggthemes)
library(VIM)
```

# Data Wrangling

```{r import_data}
exp2_rawdata <- read.csv("raw_data/quantifiers_exp2.csv", na.strings = "")
```

Take care of exclusions due to dropout and failing the attention checks.

```{r exclusions}
# remove extra rows on top
exp2_cleandata <-
  exp2_rawdata %>%
  filter(X!=1, X!=2, Status!="Survey Preview")

# number of people that dropped out mid experiment
N_dropout <-
  exp2_cleandata %>%
  filter(Finished=="False") %>%
  nrow()

# number of participants that finished the survey
N_participants <- 
  exp2_cleandata %>%
  filter(Finished=="True") %>%
  nrow()

exp2_cleandata <-
  exp2_cleandata %>%
  filter(Finished=="True", UserLanguage=="EN") %>%
  select(-X:-RecordedDate, -RecipientLastName:-example_4, -Q239...Parent.Topics:-Q239...Topics)

# number of people that failed the attention checks
N_attncheck_fail <-
  exp2_cleandata %>%
  filter(Attention.check.1_1!=70) %>%
  nrow()
```

Total of `r N_dropout + N_participants` took this survey. `r N_dropout` did not finish the survey and dropped out. The results for `r N_participants` are reported.

Tidy up the data set and transform it from wide format to long format.

```{r tidyup}
exp2_data <-
  exp2_cleandata %>%
  filter(Attention.check.1_1==70) %>% #remove those who failed attention checks
  select(-Attention.check.1_1, -Attention.check.2_1) %>% #remove attention check questions
  gather(quantifier_context_1, response, Every_Grocery_1:Never_Biked_1) %>% # make it long format
  separate(quantifier_context_1, c("quantifier", "context", "one"), sep = "_", remove=TRUE) %>%
  select(-one) %>%
  filter(response!="NA") %>%
  mutate(response=as.numeric(response))
```

Add the condition information

```{r conditions}
adverbs <- c("Never", "Rarely","Seldom", "Infreq", "Occ","Sometimes","Often","Freq","Usually","Mostly", "Always")

exp2_data$category <- "nominal"
exp2_data[exp2_data$quantifier %in% adverbs,]$category <- "adverbial"
```

Store the tidy dataframe

```{r savefile}
write_csv(exp2_data, "processed_data/exp2_data.csv")
```

Number of participants in each condition/quantifier:

```{r participantsConditions}
exp2_data %>%
  group_by(quantifier) %>%
  summarize(number = length(unique(ResponseId)))
```

# Plots

```{r}
exp2_data$quantifier <- reorder(exp2_data$quantifier, exp2_data$response, mean)

exp2_data %>%
  ggplot(aes(quantifier, response)) +
  geom_boxplot(outlier.shape=NA) +
  geom_point(size=0.1) +
  geom_jitter(size=0.1) +
  stat_summary(fun.y=median, geom="point", shape=20, size=4, color="red", fill="red") +
  theme_classic() + 
  theme(axis.text.x = element_text(size= 16, angle=45, hjust = 1, vjust = 1)) + xlab("") + ylab("")
```

## Hierarchical Cluster Analysis

We have 4 more participants per quantifier in the nominal condition. We randomly remove 4 to make both conditions mathc in number of responses for the hierarchical clustering analysis

```{r}
exp2_analysisdata <-
  exp2_data %>%
  select(-context, category) %>%
  spread(quantifier, response) 

exp2_analysisdata <- exp2_analysisdata[-sample(which(exp2_analysisdata$category=="nominal"), 4),]
```

Now we can transpose the matrix (the hclust function requires the variables to be rows!) and then run the hierarchical cluster analysis

```{r}
cluster_data <-
  exp2_analysisdata %>%
  select(-ResponseId, -category)

hclust_data <- as.data.frame(lapply(cluster_data, na.omit))

hclust_data_t <- data.frame(t(hclust_data))
colnames(hclust_data_t) <- hclust_data_t[, 1]

# Hierarchical clustering using Complete Linkage
hc1 <- hclust(
  dist(hclust_data_t, method = "manhattan"), 
  method = "complete")

# Plot the obtained dendrogram
plot(hc1, cex = 1, hang = 0.5, main= "Hierarchical Cluster Analysis", xlab="Quantifier Clusters", sub="")
```

